
---
title: "Assignment 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("analysis.R")
```

## Introduction + Summary Information
Many believe that the Civil Rights Act of 1964 marked the end of racism in the
United States. The legislation did indeed outlaw explicit discriminatory 
practices on the basis of race, color, religion, sex and national origin, 
however racism has endured and manifested itself in more quiet and insidious
form. The documentary, [13th](https://www.youtube.com/watch?v=krfcq5pF8u8), by 
*Ava DuVernay*, and book, [The New Jim Crow](https://newjimcrow.com), by 
*Michelle Alexander*, both explore the mass incarceration of African Americans, 
and the prison industrial complex as terrible manifestations of racism in the 
20th century. Factors such as over-policing people of color and the war on drugs 
have all contributed heavilytowards the disproportionate amount of people of 
color in prison systems.

This assignment serves to observe how data about county and incarcerated 
populations can highlight the patterns of inequalities present in our current 
jail and prison systems. In this project, I will be focusing upon the 
`black_jail_pop` (population of Black individuals in jail) variable(s) and 
performing analyses which highlight the disproportionate populations, by race, 
within our jail systems.

My summary analysis answers several questions & answers:

- What is the **average value** of the population of Black individuals jail 
across all the counties (in the most recent year)?

```{r avg_value_black, echo = FALSE}
avg_value_black
```

- What is the **highest value** of the population of Black individuals in jail 
across all counties (in the most recent year)?

```{r highest_value_black, echo = FALSE}
highest_value_black
```

- What county(s) and state(s) is the population of Black individuals in jail the
**highest** (in the most recent year)?

```{r county_state_highest_black, echo = FALSE}
county_state_highest_black
```

- What state(s) is the population of Black individuals in jail the **lowest** 
(in the most recent year)?

```{r state_lowest_black, echo = FALSE}
state_lowest_black
```

- How much has the total population of Black individuals in jail across all 
counties and states **changed over the last 5 years**?

```{r change_five_year, echo = FALSE}
change_five_year
```

## Trends Over Time Chart
```{r trend_chart, echo = FALSE}
trend_chart
```

```{r trend_chart_part_two, echo = FALSE}
trend_chart_part_two
```

For my **trends over time chart**, I chose to include two line graphs that show
the proportion of jail populations and the proportion of U.S populations over 
time, stratified by race. The reason I chose this chart is because it allows us 
to compare the proportions of populations by different groups of interest within
a certain population, giving us meaningful insights. Specifically, it allows us 
to compare the proportion of populations in the jail to the general U.S 
population. Through these two graphs, we can see an alarming disparity between 
their proportions. Although black populations make up a small proportion of the 
U.S population, they make up around **40%** of the population in jail systems. 
The only group that has a similar proportion in jail populations are Whites, 
which are a majority group. This highlights a problem in the United States, as 
individuals who are Black are incarcerated at a  similar rate as individuals who
are White, although they make up a significantly smaller proportion of the total
U.S population.


## Variable Comparison Chart
```{r variable_comparison_scatterplot, echo = FALSE}
variable_comparison_scatterplot
```

For my **variable comparison chart** I chose to compare two continuous variables:
`black_prison_adm_rate` (prison admission rates for Black individuals) and 
`black_prison_pop_rate` (prison population rates for Black individuals). The 
reason I selected these variables is because it provides us with an opportunity 
to understand how increasing incarceration rates for Black individuals 
ultimately creates a disproportionate population in U.S. prisons. We can see 
through this scatterplot that there is a strong positive correlation between
increasing prison admission rates for Black individuals, and an ever increasing
prison population rate for Black individuals. This is an unfortunate pattern
that continues to reinforce stereotypes and underlying forms of racism,
continuously undermining Black individuals.

# Map
```{r map, echo = FALSE}
map
```

For my **map** I chose to create a map of the United States, focusing on my own
created variable, proportions of Black individuals in jail. I divided the
`black_jail_pop` (population of Black individuals in jail) by the 
`total_jail_pop` (population of all individuals in Jail) in order to create this
value. Each state has a varying amount of blue color. A higher amount of blue 
correlates with a larger proportion of Black individuals in jail, in the most 
recent year. A lower amount of blue, or a higher amount of black, correlates 
with a smaller proportion of Black individuals in jail, in the most recent year. 
The reason I chose to use this is because it would allow us to see how different 
state shave different proportions of incarcerated Blacks. Some patterns that 
emerge from this analysis is that northern states seem to have lower proportions
of Black individuals in their jail populations, while southern states, have 
higher proportions. In addition, states in the east coast also have higher 
proportions of Black individuals in jail. This may reflect systematic racism in 
the judicial systems of these states, or racist policies that target Black 
